---
title: "HarvardX Data Science Capstone Project"
author: "Vincent"
date: "15 avril 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

# Loading and cleaning data

## Downloading and loading data

The base datafile is downloaded from Kaggle if not already available.

```{r}
# Download base dataset
DataURL <- "https://www.kaggle.com/jsphyg/weather-dataset-rattle-package/downloads/weather-dataset-rattle-package.zip/2"
DataFile <- "weather-dataset-rattle-package.zip"
if (!file.exists(DataFile)) {
  download.file(DataURL, destfile = DataFile)
  unzip(DataFile)
}
```

```{r}
rain <- read.csv("weatherAUS.csv")
str(rain)
```

The outcome variable is a factor variables with two levels, "No" coded as 1 and "Yes" coded as 2. We will recode the existing levels with a more common 0 for "No" and 1 for "Yes". This will be helpful if we later want to estimate rain probabilities, wich must obviously be between 0 and 1.

```{r}
rain$RainTomorrow2 <- as.numeric(rain$RainTomorrow == "Yes")
```



# Exploratory Data Analysys

```{r}
# Prevalence of Rain Tomorow
table(rain$RainTomorrow)
sum(rain$RainTomorrow == "No") / nrow(rain)
```
We have to take prevalence into account because in 77.5% of cases, the outcome is NO (no rain). So a dummy model always predicting no would be 77.5% accurate. we have also to consider sensitivity and specificity (ability to predict ... when ...)




# Models

## Training and testing set
```{r, warning = FALSE, message=FALSE}
library(caret)
set.seed(1971)
test_index <- createDataPartition(rain$RainTomorrow, times = 1, p = 0.2, list = FALSE)
test_set <- rain[test_index, ]
train_set <- rain[-test_index, ]
```

## Baseline model

A simple and naïve model would be to always predict the most frequent outcome. It allows us to have a baseline to evaluate more elaborate models.

```{r}
pred_naive <- rep("No", nrow(test_set))
```

### Overall Accuracy
```{r}
mean(pred_naive == test_set$RainTomorrow)
```
As expected, the overall accuracy is close to the prevalence of the most common outcome.

### Confusion Matrix

```{r}
pred_naive <- as.factor(pred_naive)
levels(pred_naive) <- levels(test_set$RainTomorrow)
table(predicted = pred_naive, actual = test_set$RainTomorrow)
```

```{r}
confusionMatrix(data = pred_naive, reference = test_set$RainTomorrow)
```

Of course our naive approach leads to a perfect sensitivity but specificity is 0. Balanced accuracy is only 0.5

## Logisitc Regression



# Results

# Conclusion

